### 2020-08-21 manuscript/summary.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Meta-analysis of the `ironic' effects of intergroup contact},
  pdfauthor={Nils Karl Reimer, University of Oxford; Nikhil Kumar Sengupta, University of Kent},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{Meta-analysis of the `ironic' effects of intergroup contact}
\author{Nils Karl Reimer, University of Oxford \and Nikhil Kumar Sengupta, University of Kent}
\date{12/08/2020}

\begin{document}
\maketitle

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{preregistered-analyses}{%
\subsection{Preregistered analyses}\label{preregistered-analyses}}

As preregistered, we ran three random-effects meta-analysis models, one
for the relationship between intergroup contact and each outcome
variable. Across 200,229 participants from 117 samples in 81 studies, we
found strong evidence for a weak association (\(r = -.07, [-.11,-.04]\))
between intergroup contact and perceived injustice, with \(>99.9\%\) of
posterior samples for the mean correlation coefficient falling below
zero. We found evidence that correlation coefficients varied across
studies (\(\tau_J = .14, [.12,.17]\)) and across samples within studies
(\(\tau_I = .09, [.06,.13]\)). Across 118,584 participants from 36
samples in 23 studies, we found some evidence for a weak association
(\(r = -.06, [-.14,.01]\)) between intergroup contact and collective
action, with \(95.0\%\) of posterior samples for the mean correlation
coefficient falling below zero. We found evidence that correlation
coefficients varied across studies (\(\tau_J = .17, [.12,.24]\)) and
across samples within studies (\(\tau_I = .09, [.06,.15]\)). Across
13,703 participants from 19 samples in 14 studies, we found some
evidence for a weak association (\(r = -.07, [-.14,-.00]\)) between
intergroup contact and policy support, with \(98.1\%\) of posterior
samples for the mean correlation coefficient falling below zero. We
found evidence that correlation coefficients varied across studies
(\(\tau_J = .10, [.06,.18]\)) and, to a lesser extent, across samples
within studies (\(\tau_I = .03, [.00,.12]\)).

As preregistered, we ran another three random-effects meta-analysis
models to estimate the relationships between the three outcome
variables. As we were not interested in the direction of these
relationships, we used cross-sectional correlation coefficients as
effect sizes for longitudinal studies. Across 111,252 participants from
24 samples in 13 studies, we found evidence for a moderate association
(\(r = .29, [.21,.37]\)) between perceived injustice and collective
action. Across 6,244 participants from 12 samples in 9 studies, we found
evidence for a moderate association (\(r = .23, [.08,.35]\)) between
perceived injustice and policy support. Across 8,558 participants from 6
samples in 3 studies, we found evidence for a moderate association
(\(r = .30, [.13,.42]\)) between collective action and policy support.

\end{document}