### 2021-03-22 manuscript/manuscript.Rmd
update posterior predictive simulations

---
title: "Meta-Analysis of the `Ironic' Effects of Intergroup Contact"
bibliography: references.bib
csl: apa.csl
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: custom-template.tex
---

```{r setup, include = FALSE}

  # Load packages
  library(tidyverse); library(tidybayes); library(numform)

  # Functions
  r_to_z <- function(r) 0.5 * log( (1 + r) / (1 - r) )
  z_to_r <- function(z) ( exp(2 * z) - 1 ) / ( exp(2 * z) + 1 )

  # knitr options
  knitr::opts_chunk$set(echo = FALSE)

```

[@pettigrew_meta-analytic_2006]

# Results

## Search results

```{=latex}
\begin{figure*}[!t]
\centering
\caption{Flow diagram illustrating the preregistered search strategy, study selection, and data collection}
\includegraphics[scale=1]{../figures/figure-1}
\label{fig:f1}
\end{figure*}
```

```{=latex}
\begin{figure*}
\centering
\caption{Overview of the relevant literature}
\includegraphics[scale=1]{../figures/figure-2}
\caption*{\textit{Note.} \textbf{A} Map of all countries included in the meta-analysis with combined sample sizes. \textbf{B} Proportion of eligible samples in each category as well as the absolute number of samples in each category.}
\label{fig:f2}
\end{figure*}
```

## Preregistered analyses

```{r, include = FALSE}

  # Summarize results
  results <- read_rds("../results/results_preregistered_analyses.rds") %>% 
    pivot_longer(
      c(r_mean, tau_ii, tau_jj),
      names_to = "parameter",
      values_to = "estimate",
      values_drop_na = TRUE
    ) %>% 
    group_by(x_var, y_var, parameter) %>% 
    summarize(
      I = f_comma(unique(I)),
      J = f_comma(unique(J)),
      N = f_comma(unique(N)),
      median = median(estimate),
      .lower = quantile(estimate, 0.025),
      .upper = quantile(estimate, 0.975),
      p_below = mean(estimate < 0),
      p_above = mean(estimate > 0)
    ) %>% 
    ungroup() %>% 
    mutate(
      across(
        c(median, .lower, .upper), 
        ~f_num(., digits = 2)
      ),
      across(
        c(p_below, p_above), 
        ~case_when(
          . < .001 ~ "$<0.1\\%$",
          . > .999 ~ "$>99.9\\%$",
          TRUE ~ paste0("$", f_num(. * 100, digits = 1), "\\%$")
        )
      ),
      text = case_when(
        parameter == "r_mean" ~ paste0("$r = ", median, ", [", .lower, ", ", .upper, "]$"),
        parameter == "tau_ii" ~ paste0("$\\tau_I = ", median, ", [", .lower, ", ", .upper, "]$"),
        parameter == "tau_jj" ~ paste0("$\\tau_J = ", median, ", [", .lower, ", ", .upper, "]$"),
        TRUE ~ paste0("$", median, ", [", .lower, ", ", .upper, "]$")
      )
    ) %>% 
    select(-median, -.lower, -.upper)

```

```{r, include = FALSE}

  # Set seed (for posterior predictive simulations)
  set.seed(9071118)

  # Simulate posterior predictions
  post_pred <- read_rds("../results/results_preregistered_analyses.rds") %>% 
    filter(x_var == "ic") %>% 
    mutate(
      r_pred = z_to_r(rnorm(n = n(), mean = mu, sd = tau_jj))
    )

  # Define functions for power calculations
  p_value <- function(r, n) 2*min(pnorm(0, r_to_z(r), 1/sqrt(n - 3)), 1-pnorm(0, r_to_z(r), 1/sqrt(n - 3)))
  p_value <- Vectorize(p_value)
  n_req   <- function(r) uniroot(function(n) ( 0.05 - p_value(r, n) ), c(3, 1e6))$root
  n_req   <- Vectorize(n_req)

  # Summarize posterior predictions
  post_pred <- post_pred %>% 
    group_by(x_var, y_var) %>% 
    summarize(
      r_pred_l80 = quantile(r_pred, 0.10),
      r_pred_u80 = quantile(r_pred, 0.90),
      r_abs = quantile(abs(r_pred), 0.20),
      n_req = n_req(r_abs)
    ) %>% 
    ungroup() %>% 
    mutate(
      across(
        c(r_pred_l80, r_pred_u80),
        ~f_num(., digits = 2)
      ),
      r_abs = paste0("$|r| > ", f_num(r_abs, digits = 3), "$"),
      n_req = f_comma(ceiling(n_req))
    )

```

As preregistered, we ran three random-effects meta-analysis models, one for each outcome variable. Figure 3 shows posterior distributions from these analyses.

```{=latex}
\begin{figure*}
\centering
\caption{Posterior distributions from the preregistered random-effect meta-analysis models}
\includegraphics[scale=1]{../figures/figure-3}
\caption*{\textit{Note.} \textbf{A} Posterior distributions for the estimated mean correlation coefficients, highlighting the proportion of posterior samples for which $r_\text{mean} < 0$. \textbf{B} Posterior predictive distributions for the estimated study-wise correlation coefficients, based on point estimates of the $\mu$ and $\tau_J$ parameters, with point estimates for the estimated correlation coefficients for all studies in the sample.}
\label{fig:f3}
\end{figure*}
```

```{r, include = FALSE}
pi_results <- results %>% filter(x_var == "ic", y_var == "pi")
```

*Perceived injustice.* Across `r unique(pi_results$N)` participants from `r unique(pi_results$I)` samples in `r unique(pi_results$J)` studies, we found strong evidence for a weak association (`r pi_results$text[pi_results$parameter == "r_mean"]`) between intergroup contact and perceived injustice, with `r pi_results$p_below[pi_results$parameter == "r_mean"]` of posterior samples for the mean correlation coefficient falling below zero. We found evidence that correlation coefficients varied across studies (`r pi_results$text[pi_results$parameter == "tau_jj"]`) and across samples within studies  (`r pi_results$text[pi_results$parameter == "tau_ii"]`). Based on these analyses, we predicted that 80\% of studies would result in correlation coefficients between `r post_pred$r_pred_l80[post_pred$y_var == "pi"]` and `r post_pred$r_pred_u80[post_pred$y_var == "pi"]` and that researchers would need sample sizes of at least `r post_pred$n_req[post_pred$y_var == "pi"]` participants to find significant associations ($\alpha = .05$, two-sided) in 80\% of their studies.^[Sample sizes are based on posterior predictions from the three models, which implied that, for 80\% of studies, the absolute correlation coefficient would be `r post_pred$r_abs[post_pred$y_var == "pi"]` for perceived injustice, `r post_pred$r_abs[post_pred$y_var == "ca"]` for collective action, and `r post_pred$r_abs[post_pred$y_var == "ps"]` for policy support.]

```{r, include = FALSE}
ca_results <- results %>% filter(x_var == "ic", y_var == "ca")
```

*Collective action.* Across `r unique(ca_results$N)` participants from `r unique(ca_results$I)` samples in `r unique(ca_results$J)` studies, we found some evidence for a weak association (`r ca_results$text[ca_results$parameter == "r_mean"]`) between intergroup contact and collective action, with `r ca_results$p_below[ca_results$parameter == "r_mean"]` of posterior samples for the mean correlation coefficient falling below zero. We found evidence that correlation coefficients varied across studies (`r ca_results$text[ca_results$parameter == "tau_jj"]`) and across samples within studies  (`r ca_results$text[ca_results$parameter == "tau_ii"]`). Based on these analyses, we predicted that 80\% of studies would result in correlation coefficients between `r post_pred$r_pred_l80[post_pred$y_var == "ca"]` and `r post_pred$r_pred_u80[post_pred$y_var == "ca"]` and that researchers would need sample sizes of at least `r post_pred$n_req[post_pred$y_var == "ca"]` participants to find significant associations ($\alpha = .05$, two-sided) in 80\% of their studies.

```{r, include = FALSE}
ps_results <- results %>% filter(x_var == "ic", y_var == "ps")
```

*Policy support.* Across `r unique(ps_results$N)` participants from `r unique(ps_results$I)` samples in `r unique(ps_results$J)` studies, we found some evidence for a weak association (`r ps_results$text[ps_results$parameter == "r_mean"]`) between intergroup contact and policy support, with `r ps_results$p_below[ps_results$parameter == "r_mean"]` of posterior samples for the mean correlation coefficient falling below zero. We found evidence that correlation coefficients varied across studies (`r ps_results$text[ps_results$parameter == "tau_jj"]`) and, to a lesser extent, across samples within studies  (`r ps_results$text[ps_results$parameter == "tau_ii"]`). Based on these analyses, we predicted that 80\% of studies would result in correlation coefficients between `r post_pred$r_pred_l80[post_pred$y_var == "ps"]` and `r post_pred$r_pred_u80[post_pred$y_var == "ps"]` and that researchers would need sample sizes of at least `r post_pred$n_req[post_pred$y_var == "ps1"]` participants to find significant associations ($\alpha = .05$, two-sided) in 80\% of their studies.

```{r, include = FALSE}
pi_ca_results <- results %>% filter(x_var == "pi", y_var == "ca", parameter == "r_mean")
pi_ps_results <- results %>% filter(x_var == "pi", y_var == "ps", parameter == "r_mean")
ca_ps_results <- results %>% filter(x_var == "ca", y_var == "ps", parameter == "r_mean")
```

As preregistered, we ran another three random-effects meta-analysis models to estimate the relationships between the three outcome variables. As we were not interested in the direction of these relationships, we used cross-sectional correlation coefficients as effect sizes for longitudinal studies. Across `r pi_ca_results$N` participants from `r pi_ca_results$I` samples in `r pi_ca_results$J` studies, we found evidence for a moderate association (`r pi_ca_results$text`) between perceived injustice and collective action. Across `r pi_ps_results$N` participants from `r pi_ps_results$I` samples in `r pi_ps_results$J` studies, we found evidence for a moderate association (`r pi_ps_results$text`) between perceived injustice and policy support. Across `r ca_ps_results$N` participants from `r ca_ps_results$I` samples in `r ca_ps_results$J` studies, we found evidence for a moderate association (`r ca_ps_results$text`) between collective action and policy support.

### Robustness checks

```{r, include = FALSE}

  # Summarize prior choice results
  d_priors <- read_rds("../results/results_robustness_checks_priors.rds") %>% 
    group_by(y_var, prior) %>% 
    median_qi(d_r_mean) %>% 
    mutate(
      across(c(d_r_mean, .lower, .upper), ~f_num(., digits = 2)),
      text = paste0("$\\Delta r = ", d_r_mean, ", [", .lower, ", ", .upper, "]$")
    )

```

```{r, include = FALSE}

  # Calculate Mean Absolute Deviation (MAD)
  mad_loo <- read_rds("../results/results_robustness_checks_loo.rds") %>% 
    group_by(.draw, y_var) %>% 
    summarize(mad_loo = mean(ae_loo)) %>% 
    group_by(y_var) %>% 
    median_qi(mad_loo) %>% 
    mutate(
      across(c(mad_loo, .lower, .upper), ~f_num(., digits = 2)),
      text = paste0("$\\textit{MAD} = ", mad_loo, ", [", .lower, ", ", .upper, "]$")
    )

  # Leave out most influential studies
  d_loo <- read_rds("../results/results_robustness_checks_loo.rds") %>% 
    group_by(y_var, id) %>%
    median_qi(e_loo) %>% 
    top_n(n = 1, wt = abs(e_loo)) %>% 
    mutate(
      across(c(e_loo, .lower, .upper), ~f_num(., digits = 2)),
      text = paste0("$\\Delta r = ", e_loo, ", [", .lower, ", ", .upper, "]$")
    )

```

First, we assessed to what extent our findings were sensitive to choosing narrower, $\mu \sim \text{Normal}(0, 0.1)$, or wider, $\mu \sim \text{Normal}(0, 1)$, prior distributions. Choosing narrower or wider prior distribution did not affect mean effect size estimates for perceived injustice (`r d_priors$text[d_priors$y_var == "pi" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "pi" & d_priors$prior == "N(0, 1.0)"]`), collective action (`r d_priors$text[d_priors$y_var == "ca" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "ca" & d_priors$prior == "N(0, 1.0)"]`), and policy support (`r d_priors$text[d_priors$y_var == "ps" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "ps" & d_priors$prior == "N(0, 1.0)"]`). Second, we assessed to what extent our findings were sensitive to including or excluding influential studies by repeating the preregistered analyses $J$ times while leaving out one of $J$ studies each time and by calculating the mean absolute difference (*MAD*) for the estimated mean effect size across left-out studies. For perceived injustice (`r mad_loo$text[mad_loo$y_var == "pi"]`), collective action (`r mad_loo$text[mad_loo$y_var == "ca"]`), and policy support (`r mad_loo$text[mad_loo$y_var == "ps"]`), the *MAD* was small. Leaving out the most influential study, for example, did not change estimates of the mean effect size for the three outcomes (`r d_loo$text[d_loo$y_var == "pi"]`; `r d_loo$text[d_loo$y_var == "ca"]`; `r d_loo$text[d_loo$y_var == "ps"]`). Together, these analyses showed that our findings were robust to choosing different prior distributions and to excluding influential studies.

# References

\begingroup

\noindent
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs"></div>

\endgroup