### 2021-04-13 manuscript/supplemental-materials.Rmd

---
title: "Supplemental Materials"
bibliography: references.bib
csl: apa.csl
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: custom-template.tex
---

```{r setup, include = FALSE}

  # Load packages
  library(tidyverse); library(tidybayes); library(numform)

  # Functions
  r_to_z <- function(r) 0.5 * log( (1 + r) / (1 - r) )
  z_to_r <- function(z) ( exp(2 * z) - 1 ) / ( exp(2 * z) + 1 )

  # knitr options
  knitr::opts_chunk$set(echo = FALSE)

```

# Robustness Checks

```{r, include = FALSE}

  # Summarize prior choice results
  d_priors <- read_rds("../results/results_robustness_checks_priors.rds") %>% 
    group_by(y_var, prior) %>% 
    median_qi(d_r_mean) %>% 
    mutate(
      across(c(d_r_mean, .lower, .upper), ~f_num(., digits = 2)),
      text = paste0("$\\Delta r = ", d_r_mean, ", [", .lower, ", ", .upper, "]$")
    )

```

```{r, include = FALSE}

  # Calculate Mean Absolute Deviation (MAD)
  mad_loo <- read_rds("../results/results_robustness_checks_loo.rds") %>% 
    group_by(.draw, y_var) %>% 
    summarize(mad_loo = mean(ae_loo)) %>% 
    group_by(y_var) %>% 
    median_qi(mad_loo) %>% 
    mutate(
      across(c(mad_loo, .lower, .upper), ~f_num(., digits = 2)),
      text = paste0("$\\textit{MAD} = ", mad_loo, ", [", .lower, ", ", .upper, "]$")
    )

  # Leave out most influential studies
  d_loo <- read_rds("../results/results_robustness_checks_loo.rds") %>% 
    group_by(y_var, id) %>%
    median_qi(e_loo) %>% 
    top_n(n = 1, wt = abs(e_loo)) %>% 
    mutate(
      across(c(e_loo, .lower, .upper), ~f_num(., digits = 2)),
      text = paste0("$\\Delta r = ", e_loo, ", [", .lower, ", ", .upper, "]$")
    )

```

As preregistered, we conducted two kinds of robustness checks. First, we assessed to what extent our findings were sensitive to choosing narrower, $\mu \sim \text{Normal}(0, 0.1)$, or wider, $\mu \sim \text{Normal}(0, 1)$, prior distributions. Choosing narrower or wider prior distribution did not affect mean effect size estimates for perceived injustice (`r d_priors$text[d_priors$y_var == "pi" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "pi" & d_priors$prior == "N(0, 1.0)"]`), collective action (`r d_priors$text[d_priors$y_var == "ca" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "ca" & d_priors$prior == "N(0, 1.0)"]`), and policy support (`r d_priors$text[d_priors$y_var == "ps" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "ps" & d_priors$prior == "N(0, 1.0)"]`). Second, we assessed to what extent our findings were sensitive to including or excluding influential studies by repeating the preregistered analyses $J$ times while leaving out one of $J$ studies each time and by calculating the mean absolute difference (*MAD*) for the estimated mean effect size across left-out studies. For perceived injustice (`r mad_loo$text[mad_loo$y_var == "pi"]`), collective action (`r mad_loo$text[mad_loo$y_var == "ca"]`), and policy support (`r mad_loo$text[mad_loo$y_var == "ps"]`), the *MAD* was small. Leaving out the most influential study, for example, did not change estimates of the mean effect size for the three outcomes (`r d_loo$text[d_loo$y_var == "pi"]`; `r d_loo$text[d_loo$y_var == "ca"]`; `r d_loo$text[d_loo$y_var == "ps"]`). Together, these analyses showed that our findings were robust to choosing different prior distributions and to excluding influential studies.

# References

\begingroup

\noindent
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs"></div>

\endgroup